---
title: "Two-proportion inference and linear prediction using a documented random sample of neonatal hospital
records"
author: 
  - "Group Members:"
  - "Ikenna Ezike"
  - "Milly Tesfazgy"
  - "Million Seghid"
date: "2026-09-25"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

##### Introduction

Purpose and guiding questions
We will study neonates diagnosed with meningitis and admitted to the
neonatal intensive care units at Hiwot Fana
Comprehensive Specialized University Hospital and Jugal General Hospital
in Harar, Ethiopia, from 1 October 2020
through 31 October 2024. The finite target population is the eligible
admission records in those two hospitals
during that interval, not all babies in Ethiopia or future admissions
[1]. The complete hospital registry is not
supplied; the public file contains a sample of records [1, 2].
Inference: Among that population, is the proportion with a Poor
prognosis higher in the original Gestational age
category \<37 weeks than in the category \>=37 weeks? Prediction: Among
comparable eligible neonates, how well
does the original Gestational age of neonate (week) predict Birth weight
of neonates (gram)? The questions use the
same sampled records but distinct response variables.

```{r}
library(haven)
my_data <- read_sav("C:/Users/Ikemma/Downloads/new data 11111111.sav")

# Preterm: gestational age < 37 AND outcome = 2
preTerm <- my_data[
  my_data$`@5.Gestationalageofneonateweek` < 37 &
    my_data$outcome == 2,
]

# Term: gestational age >= 37 AND outcome = 2
postTerm <- my_data[
  my_data$`@5.Gestationalageofneonateweek` >= 37 &
    my_data$outcome == 2,
]

# All observations with outcome = 2
poorTreatment <- my_data[
  my_data$outcome == 2,
]


preTerm_number = nrow(postTerm)
postTerm_number = nrow(preTerm)
poorTreatmentToal = nrow(poorTreatment)

#Pre term proportion p1
preTerm_number / poorTreatmentToal

#Post term proportion p2
postTerm_number / poorTreatmentToal



```

\

\

\$\$\$\$
